draft 1

### 2016-06-21 finalwriteup/draft.rmd
finished blended model method

ANOVA - Anaylis of Variance. A statistical test that can be applied to find whether there is a significant difference between the means of more than two groups.
B2B - Business to Business 
CBR - Case based reasoning. A method of estimating by using the results of similar cases.
CRM - Customer Relationship Management. Refers to a common type of software used by businesses to record client and project details as well as employee timesheet records. Generally accessable to each employee in a company.
DAG - directed acyclic graph. A graph consisting of vertices connected by edges, where each edge directs a vertex to another vertex. The arrangement of the edges is such that a vertex cannot follow a path that loops back to itself.
IT - Information technology
BIM - Building Information Modelling. A digital representation of physical characteristics of an object or structure.
ROC - Reciever Operating Characteristic
AUC - Area under the curve. In this project, the area refers to the area under an ROC curve
TPR - True positive rate
FPR - False positive rate
TP - true positive
FP - false positive
DA - Discriminant analysis
CLV - Customer Lifetime Value
IQR - Interquartile range
ESS - error sum of squares
RMSE - root mean squared error

@@

In order to compare which models performed significantly better than others, an adequate sample size of results statistics was required. Multiple models could be made by using different data in the training vs. testing sets, each providing a resulting test statistic (RMSE or AUC). 5 fold cross validation was used, which created 5 models, then the division of 5 folds was repeated using a different random seed to create another set of 5 training/testing data sets. Initially, 20 models of each method were created in this fashion. Then a two-sample power calculation was run using the two sets of 20 results to determine the sample size to achieve a statistical power of 0.8. Once the number of required samples was determined, more models were built via 5 fold cross validation to obtain additional results statistics. The highest performing models were advanced to the next stage where various methods of combining, or blending, model results were tested.

@@

- simple averaigng, log reg, FWLS, boosted trees random forests
- repeat 20 times - power - create more samples
- carry all forward for bottom line translation anyway

@@

Although predictive power of the final model is important, for the method to be integrated into a company's decisions, the affect on a business' bottom line was tested. This was done using a profit curve - a chart that plots the amount of profit the company earns on the y-axis vs. the probability threshold on the x-axis.

>>>>

ANOVA - Anaylis of Variance. A statistical test that can be applied to find whether there is a significant difference between the means of more than two groups.
B2B - Business to Business 
CBR - Case based reasoning. A method of estimating by using the results of similar cases.
CRM - Customer Relationship Management. Refers to a common type of software used by businesses to record client and project details as well as employee timesheet records. Generally accessable to each employee in a company.
DAG - directed acyclic graph. A graph consisting of vertices connected by edges, where each edge directs a vertex to another vertex. The arrangement of the edges is such that a vertex cannot follow a path that loops back to itself.
IT - Information technology
BIM - Building Information Modelling. A digital representation of physical characteristics of an object or structure.
ROC - Reciever Operating Characteristic
AUC - Area under the curve. In this project, the area refers to the area under an ROC curve
TPR - True positive rate
FPR - False positive rate
TP - true positive
FP - false positive
DA - Discriminant analysis
CLV - Customer Lifetime Value
IQR - Interquartile range
ESS - error sum of squares
RMSE - root mean squared error
FWLS - feature weighted linear stacking

@@

In order to compare which models performed significantly better than others, an adequate sample size of results statistics was required. Multiple models could be made by using different data in the training vs. testing sets, each providing a resulting test statistic (RMSE or AUC). 5 fold cross validation was used, which created 5 models, then the division of 5 folds was repeated using a different random seed to create another set of 5 training/testing data sets. Initially, 20 models of each method were created in this fashion. Then a two-sample power calculation was run using the two sets of 20 results to determine the sample size to achieve a statistical power of 0.8. Once the number of required samples was determined, more models were built via 5 fold cross validation to obtain additional results statistics. The highest performing models were advanced to the next stage where various methods of combining, or blending, methods were tested.

@@

Several research groups involved in the high profile data science competition, Netflix Prize, developed sophisticated methods of model stacking, otherwise known as ensemble methods or model blending [@Sill2009]. In this research the term 'blending' is used to avoid confusion between ensemble tree methods. The idea behind model blending is to combine predictions from several methods with different theoretical foundations so that the strengths of each method can be combined. Historically, this was called model averaging in statistics and since the 1990's research showed that averaging results from different methods provided better predictive accuracy than any single model [@Madigan1994]. New methods of model blending were developed for the Netflix Prize that applied sophisiticated machine learning techniques to big, messy data. The prize-winning solution was a complex blend of sub-blends that incorporated results of individual models as variables along with original variables. When original variables are used in blending they are called meta-features [@Sill2009]. The multiple layers of blending gave incrememntal improvements in predictive accuracy, but the computational cost of the complex layers did not justify benefits for Netflix in practice and a simplified model was adopted. Therefore, in this project a single layer of blending was trialled.

Five blending methods, ranging from simple to complex were tested on the top performing individual models. These included simple averaging of the individual model results, building a regression model using the individual model results only, feature weighted linear stacking (FWLS), random forests, and boosted trees. The first two models are simply the average or weighted average of the three individual models. The last three methods take advantage of potentially meaningful interactions between an individual model's results and meta-features (the original variables). For example, if the regression predicted profitability better than the other models for Water projects, the model would take advantage of this interaction. Regression predictions would be weighted higher than the other model predictions for Water projects.

A simplified explanation of FWLS is as a linear regression where meta-features as well as model predictions from individual models are included as variables. Then, each meta-feature is interacted with each column of model results [@Sill2009]. A linear regression is performed to weight each term's contribution to the final prediction. Random forests and boosted trees are also used to blend meta-features with output from several models. Feature interaction is performed passively due to the nature of how trees are built. A split in a node is conditional upon the preceding split based on another variable and so on.

The five blended models were built on training and testing sets multiple times in a similar fashion to the individual models. A single training set for a blended model must be built from the test results of the individual models. Therefore, five fold cross validation was performed using each individual model to create a full dataset of test results. This was then added to the original dataset, and training and testing data could be partitioned to test the blended models. A maximum of five blended models were created from a complete set of test results, then a new set of test results were created for the next five blended models. This ensured a good mix of training and test data was employed.

To begin, 20 models of each blending method were created. Then, a two-sample power calculation was performed to achieve statistical power of 0.8. The required number of samples (models) were then developed so that the highest performing blended model could be tested against the others for statistical significance. The next step was to assess the blended models' performance on overall profits for the company. This analysis differed from predicitve model accuracy, so several blended models as well as individual models were carried forward for this analysis.

@@

Although predictive power of the final model was important, for the method to be integrated into a company's decisions, the affect on a business' bottom line was tested. This was done using a profit curve - a chart that plots the amount of profit the company earns on the y-axis vs. the probability threshold on the x-axis.
